---
title: "STAN Tutorial"
author: "Seongwon Im"
date: "2/18/2021"
output: 
  html_document:
    theme: yeti
---
## rstan

Description of 
what is stan? 
why is stan?

```{r, message = FALSE}
# Import packages
library(rstan)
library(rstanarm)
```

### 1. Data

Description of data

```{r, fig.align='center'}
# Built in dataset: kidiq
mom.iq <- kidiq$mom_iq # Predictor
child.iq <- kidiq$kid_score # Response

# Plot
plot(mom.iq, child.iq, xlab = "Mom IQ", ylab = "Child IQ")
```

## 2. Preparing data

Preparing the data? What is the input?
```{r}
stan_data = list(N = nrow(kidiq), # Number of rows of the dataset
                 x = mom.iq, # Predictor variable
                 y = child.iq) # Response variable

# Design matrix for multiple covariates
# Parameters for prior distributions
```

### 3. Writing a stan code

Basic Format
```{r}
write("
data { // This is where you format the input data (Known)
}

parameters { // This is where you state your parameters (Unknown)
}

transformed parameters{ // This is where you transform your parameters
}                       // Good for hierarchical modeling

model { // This is where you specify distributions.
}

generated quantities { // Good for prediction
}"

, "name.stan") # Name the stan file
```
Explanation on each block


### Simple linear regression 
$$Y \sim N(X \beta, \sigma^2 I)$$
Explanation of simple linear regression

Known: $\beta_0$, $\beta_1$, $\sigma^2$

Unknown: $Y$, $X$
```{r}
# Create a stan file.
write("
data { 
 int <lower = 1> N; // Sample size with restriction
 vector[N] x; // Predictor variable
 vector[N] y; // Response variable
}

parameters { 
 real beta0; // Intercept
 real beta1; // Slope
 real <lower = 0> sigma; // Error
}

transformed parameters{
}

model {
 //beta0 ~ normal(10, 0.1); // if want prior for the intercept
 //beta1 ~ normal(1, 0.1); // if want prior for the slope
 y ~ normal(beta0 + x * beta1, sigma);
}

generated quantities {
}"

, "stanSLR.stan") # Name the stan file stanSLR.stan
```

### 4. Fit the model
```{r, message = FALSE}
# fit with lm() to see what to expect
lmfit = lm(child.iq ~ mom.iq)
c(lmfit$coefficients, sigma(lmfit))
```

Explain the result

```{r, results='hide'}
# Fit the model
stanfit <- stan(file = "stanSLR.stan", # Stan file
                data = stan_data, # Data
                warmup = 500, # Number of iteration to burn-in
                iter = 1500, # Number of iteration
                chains = 4) # Number of chains to run
```
What to include in the stan model

### 5. Analysis
```{r, fig.align='center'}
# See the fit
stanfit

# Plot the fit
plot(mom.iq, child.iq, xlab = "Mom IQ", ylab = "Child IQ") # Plot the data
abline(mean(extract(stanfit)$beta0), mean(extract(stanfit)$beta1), 
       col = 4, lw = 2) # Include the line
```


```{r, warning = FALSE, fig.align='center'}
# More plots
traceplot(stanfit)
stan_dens(stanfit)
stan_hist(stanfit)
```



```{r, fig.align='center'}
# Checking convergence
#shinystan::launch_shinystan(stanfit)
```

&nbsp;

## Reference

* Coding Club. (n.d.). Intro to Stan. Retrieved February 15, 2021, from https://ourcodingclub.github.io/tutorials/stan-intro/

* Gabry, J. G., & Goodrich, B. G. (2020, July 20). Estimating 
Regularized Linear Models with rstanarm. Estimating Regularized Linear Models with Rstanarm. 

* https://mc-stan.org/rstanarm/articles/lm.htmlLai, M. (2019, December 13). Course Handouts for Bayesian Data Analysis Class. Course Handouts for Bayesian Data Analysis Class. https://bookdown.org/marklhc/notes_bookdown/
